---
title: 'Immune Modulation Effect of VEGF-A Inhibition with Bevacizumab Combined with
  PD-L1 Blockade with Avelumab for Recurrent Glioblastoma '
output:
  html_document: default
  pdf_document: default
date: '2022-06-21'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(tidyr)
library(patchwork)
library(janitor)
library(survival)
library(survminer)
library(DT)
```

### Load clinical data cleaned
```{r, warning=F, message=F}
load("../data/data.RData")
```

### Survival analysis patient received bevacizumab vs not received bevacizumab
```{r, warning=F, message=F}
censored_at <- 540 # 18 months

bldf <- clinical_data%>%
  group_by(subject_id)%>%
  arrange(trt_day_from_admission)%>%
  filter(row_number() == 1)%>%
  mutate(
    #----death
         censored_death_day_from_admission = ifelse(is.na(death_day_from_admission) | (death_day_from_admission > censored_at), censored_at, death_day_from_admission))%>%
              ungroup()%>%
  dplyr::select(subject_id, ever_bev, censored_death_day_from_admission, death)%>%
  ungroup()


# death outcome
fit.one <- survfit(Surv(censored_death_day_from_admission, death == "died") ~ ever_bev, data = bldf, conf.type = "log-log")

```

### Fig 3A
```{r, warning=F, message=F}
survminer::ggsurvplot(fit.one, data = bldf, conf.int = T, 
                      xlab = "Time from Admission",  break.time.by = 30,
                      ncensor.plot = F, risk.table = T)
```

### Univariable analysis Olink Cytokine

#### Difference in expression patient received bevacizumab vs not received bevacizumab
```{r, warning=F, message=F}
data_use <- cbind(meta_df, olink_df)
data_use$trt_group <- factor(data_use$trt_group, levels = c("baseline", "ave", "combo"))

#------parametric
design <- model.matrix( ~ ever_bev, data = data_use )  #  bev  ever_bev
fit <- limma::lmFit(t(olink_df), design, method = "robust")
fit <- limma::contrasts.fit(fit, coefficients = ncol(design))
efit <- limma::eBayes(fit, robust = TRUE)
#------nonparametric
rr <- lapply(colnames(olink_df), function(x){
  
  temp <- data_use%>%
    dplyr::select(c(x, "bev", "ever_bev"))%>%
    mutate(bev = factor(bev))
  fit <- wilcox.test(get(x) ~ ever_bev, data = temp, paired = F)

  Sum <- temp%>%
    group_by(bev)%>%
    summarize(Median = median(get(x), na.rm = T))

  re <- c(x, fit$p.value, Sum[1 , 2], Sum[2 , 2])
  names(re) <- c("marker", "wilcox_pval", "median_bev_no", "median_bev_yes")
  
  unlist(re)
})

rr <- rr%>%
  do.call(what = "rbind")%>%
  data.frame()%>%
  group_by(marker)%>%
  mutate_all(as.numeric)%>%
  ungroup()%>%
  mutate(adj_wilcox_pval = p.adjust(wilcox_pval, method = "BH")) 


```


#### Difference in expression patient patient at baseline, avelumab, and combo(avelumab plus bevacizumab)
```{r, warning=F, message=F}
rr <- lapply(colnames(olink_df), function(x){

  fit <- aov(get(x) ~ trt_group, data = data_use%>%
               dplyr::select(x, "trt_group", "subject_id"))
  anova_pval <- summary(fit)[[1]][["Pr(>F)"]][1]
  fit_post <- try(summary(glht(fit, linfct = mcp(trt_group = "Tukey"))))
  fit1 <- kruskal.test(get(x) ~ trt_group, data = data_use)
  fit1_post<- pairwise.wilcox.test(unlist(data_use[,x]), unlist(data_use[,"trt_group"]), p.adj="none")
  mediandf <- data_use%>%
    group_by(trt_group)%>%
    summarize(median(get(x)))
  re <- c(x, anova_pval, fit_post$test$coefficients, fit_post$test$pvalues, 
          fit1$p.value, fit1_post$p.value[1,1], fit1_post$p.value[2,1], fit1_post$p.value[2,2],
          mediandf[, 2]%>% unlist())
  names(re) <- c("marker", "pvalue_anova", paste0("coeff_p_", names(fit_post$test$coefficients)), paste0("pvalue_p_", names(fit_post$test$coefficients)), 
                 "pvalue_kruskal", "pvalue_np_baseline..ave", "pvalue_np_baseline..combo", "pvalue_np_combo..ave",
                 paste0("median_", mediandf[, 1]%>% unlist()%>% as.character()))
  
  re
})



rr <- do.call(rbind, rr)%>%
  data.frame()%>%
  group_by(marker)%>%
  mutate_all(as.numeric)%>%
  ungroup()%>%
  mutate(adj_pval_anova = p.adjust(pvalue_anova, method = "BH"),
         adj_pvalue_kruskal = p.adjust(pvalue_kruskal, method = "BH"))

colnames(rr) <- gsub("[\\.]+", "2", colnames(rr))

```

### Table S2
```{r, warning=F, message=F}
rr%>%
  filter(adj_pval_anova <= 0.05)%>%
  DT::datatable(options = list(scrollX = TRUE))%>%
  DT::formatRound(columns = colnames(rr)[-1],
                  digits = 3)
  
```

### Fig 3B
```{r, warning=F, message=F, fig.width=12}
plot.ls <- lapply(
  rr%>%filter(adj_pval_anova <= 0.05)%$%marker, function(x){
  trt_group_td%>%
    dplyr::select(trt_group, x, subject_id)%>%
    ggplot()+
    geom_boxplot(aes(trt_group, get(x), fill = trt_group), alpha = 0.7)+
    geom_point(aes(trt_group, get(x), color = trt_group), size = 3, position = position_dodge(width = 0.7))+
    scale_color_manual(values = c("black", "black", "black"))+
    labs(title = x, x = "", y = "")+
    theme_bw()
})
wrap_plots(plot.ls, nrow = 2) + plot_layout(guides = "collect")
```

### Univariable analysis Cytof Cellsubtype frequency

#### Difference in expression patient received bevacizumab vs not received bevacizumab child population 
```{r, warning=F, message=F}
filt_sample <- apply(cytof_child_df, 1, function(x) sum(is.na(x)) != length(x))
data_use <- cbind(meta_df, cytof_child_df)[filt_sample, ]
data_use$trt_group <- factor(data_use$trt_group, levels = c("baseline", "ave", "combo"))

#------parametric
design <- model.matrix( ~ ever_bev, data = data_use )  #  bev  ever_bev
fit <- limma::lmFit(t(cytof_child_df)[ , filt_sample], design, method = "robust")
fit <- limma::contrasts.fit(fit, coefficients = ncol(design))
efit <- limma::eBayes(fit, robust = TRUE)
#------nonparametric
rr <- lapply(colnames(cytof_child_df), function(x){
  
  temp <- data_use%>%
    dplyr::select(c(x, "bev", "ever_bev"))%>%
    mutate(bev = factor(bev))
  fit <- wilcox.test(get(x) ~ ever_bev, data = temp, paired = F)

  Sum <- temp%>%
    group_by(bev)%>%
    summarize(Median = median(get(x), na.rm = T))

  re <- c(x, fit$p.value, Sum[1 , 2], Sum[2 , 2])
  names(re) <- c("marker", "wilcox_pval", "median_bev_no", "median_bev_yes")
  
  unlist(re)
})

rr <- rr%>%
  do.call(what = "rbind")%>%
  data.frame()%>%
  group_by(marker)%>%
  mutate_all(as.numeric)%>%
  ungroup()%>%
  mutate(adj_wilcox_pval = p.adjust(wilcox_pval, method = "BH")) 

```

### Fig 3C
```{r, warning=F, message=F}
plot.ls <- lapply(rr%>%
  filter((wilcox_pval <= 0.05))%$%marker, function(x){
  data_use%>%
    dplyr::select(ever_bev, x)%>%
    ggplot()+
    geom_boxplot(aes(ever_bev, get(x), fill = ever_bev,
                     ), alpha = 0.7)+
    geom_point(aes(ever_bev, get(x), color = ever_bev , 
                   ), size = 3, position = position_dodge(width = 0.7))+
    scale_color_manual(values = c("black", "black"))+
    labs(title = x, x = "", y = "")+
    theme_bw()
})

wrap_plots(plot.ls, nrow = 1) + plot_layout(guides = "collect")
```

#### Difference in expression patient patient at baseline, avelumab, and combo(avelumab plus bevacizumab) child population 
```{r, warning=F, message=F}

rr <- lapply(colnames(cytof_child_df), function(x){

  fit <- aov(get(x) ~ trt_group, data = data_use%>%
               dplyr::select(x, "trt_group", "subject_id"))
  anova_pval <- summary(fit)[[1]][["Pr(>F)"]][1]
  fit_post <- try(summary(glht(fit, linfct = mcp(trt_group = "Tukey"))))
  fit1 <- kruskal.test(get(x) ~ trt_group, data = data_use)
  fit1_post<- pairwise.wilcox.test(unlist(data_use[,x]), unlist(data_use[,"trt_group"]), p.adj="none")
  mediandf <- data_use%>%
    group_by(trt_group)%>%
    summarize(median(get(x)))
  re <- c(x, anova_pval, fit_post$test$coefficients, fit_post$test$pvalues, 
          fit1$p.value, fit1_post$p.value[1,1], fit1_post$p.value[2,1], fit1_post$p.value[2,2],
          mediandf[, 2]%>% unlist())
  names(re) <- c("marker", "pvalue_anova", paste0("coeff_p_", names(fit_post$test$coefficients)), paste0("pvalue_p_", names(fit_post$test$coefficients)), 
                 "pvalue_kruskal", "pvalue_np_baseline..ave", "pvalue_np_baseline..combo", "pvalue_np_combo..ave",
                 paste0("median_", mediandf[, 1]%>% unlist()%>% as.character()))
  
  re
})


rr <- do.call(rbind, rr)%>%
  data.frame()%>%
  group_by(marker)%>%
  mutate_all(as.numeric)%>%
  ungroup()%>%
  mutate(adj_pval_anova = p.adjust(pvalue_anova, method = "BH"),
         adj_pvalue_kruskal = p.adjust(pvalue_kruskal, method = "BH"))

colnames(rr) <- gsub("[\\.]+", "2", colnames(rr))

```


#### Difference in expression patient received bevacizumab vs not received bevacizumab parent population 
```{r, warning=F, message=F}
filt_sample <- apply(cytof_parent_df, 1, function(x) sum(is.na(x)) != length(x))
data_use <- cbind(meta_df, cytof_parent_df)[filt_sample, ]
data_use$trt_group <- factor(data_use$trt_group, levels = c("baseline", "ave", "combo"))

#------parametric
design <- model.matrix( ~ ever_bev, data = data_use )  #  bev  ever_bev
fit <- limma::lmFit(t(cytof_parent_df)[ , filt_sample], design, method = "robust")
fit <- limma::contrasts.fit(fit, coefficients = ncol(design))
efit <- limma::eBayes(fit, robust = TRUE)



#------nonparametric
rr <- lapply(colnames(cytof_parent_df), function(x){
  
  temp <- data_use%>%
    dplyr::select(c(x, "bev", "ever_bev"))%>%
    mutate(bev = factor(bev))
  fit <- wilcox.test(get(x) ~ ever_bev, data = temp, paired = F)

  Sum <- temp%>%
    group_by(bev)%>%
    summarize(Median = median(get(x), na.rm = T))

  re <- c(x, fit$p.value, Sum[1 , 2], Sum[2 , 2])
  names(re) <- c("marker", "wilcox_pval", "median_bev_no", "median_bev_yes")
  
  unlist(re)
})

rr <- rr%>%
  do.call(what = "rbind")%>%
  data.frame()%>%
  group_by(marker)%>%
  mutate_all(as.numeric)%>%
  ungroup()%>%
  mutate(adj_wilcox_pval = p.adjust(wilcox_pval, method = "BH")) 

```

#### Difference in expression patient patient at baseline, avelumab, and combo(avelumab plus bevacizumab) parent population 
```{r, warning=F, message=F}
rr <- lapply(colnames(cytof_parent_df), function(x){

  fit <- aov(get(x) ~ trt_group, data = data_use%>%
               dplyr::select(x, "trt_group", "subject_id"))
  anova_pval <- summary(fit)[[1]][["Pr(>F)"]][1]
  fit_post <- try(summary(glht(fit, linfct = mcp(trt_group = "Tukey"))))
  fit1 <- kruskal.test(get(x) ~ trt_group, data = data_use)
  fit1_post<- pairwise.wilcox.test(unlist(data_use[,x]), unlist(data_use[,"trt_group"]), p.adj="none")
  mediandf <- data_use%>%
    group_by(trt_group)%>%
    summarize(median(get(x)))
  re <- c(x, anova_pval, fit_post$test$coefficients, fit_post$test$pvalues, 
          fit1$p.value, fit1_post$p.value[1,1], fit1_post$p.value[2,1], fit1_post$p.value[2,2],
          mediandf[, 2]%>% unlist())
  names(re) <- c("marker", "pvalue_anova", paste0("coeff_p_", names(fit_post$test$coefficients)), paste0("pvalue_p_", names(fit_post$test$coefficients)), 
                 "pvalue_kruskal", "pvalue_np_baseline..ave", "pvalue_np_baseline..combo", "pvalue_np_combo..ave",
                 paste0("median_", mediandf[, 1]%>% unlist()%>% as.character()))
  
  re
})


rr <- do.call(rbind, rr)%>%
  data.frame()%>%
  group_by(marker)%>%
  mutate_all(as.numeric)%>%
  ungroup()%>%
  mutate(adj_pval_anova = p.adjust(pvalue_anova, method = "BH"),
         adj_pvalue_kruskal = p.adjust(pvalue_kruskal, method = "BH"))

colnames(rr) <- gsub("[\\.]+", "2", colnames(rr))

```

```{r, echo=F, eval=F}
rmarkdown::render(input = './main.Rmd', output_dir = '../', clean = TRUE, rmarkdown::html_document(pandoc_args = "--self-contained"))
```

